RNAseq aCRY pCRY data dive

### 2024-04-21 2_A_dive_aCRY_pCRY/2_A_dive_aCRY_pCRY.Rmd
include pCRY figures

# 3. Results
## - Plot Counts

```{r example_counts, fig.show="hold", out.width="50%", eval=FALSE, include=TRUE, echo=FALSE}

# search genes
anno[str_detect(anno[["geneSymbol"]],"CRY"),1:9] %>% kable()

anno[str_detect(anno[["geneSymbol"]],"ROC"),1:9] %>% kable()
anno[str_detect(anno[["previousIdentifiers"]],"CRY"),1:9] %>% kable()

# plap6
PLAP6 <- "Cre03.g188700"
d <- plotCounts(dds,gene = PLAP6, intgroup= c("treatment","genotype","condition"), returnData=TRUE)

resultsNames(dds)
res <- results(dds,contrast = c("genotype","cia5","WT"))
res["Cre03.g188700",]
ggplot(d, aes(x = genotype, y = count, fill=genotype, color=genotype)) +
    geom_boxplot(color="black") +
    geom_point(shape=21,color="black",aes(fill=genotype),position=position_dodge(width=0.75), alpha=1) +
   # scale_fill_manual(values=cols[c(1,5)]) +
  #  scale_color_manual(values=cols[c(1,5)]) +
   scale_y_continuous(trans = "log2") +
  scale_fill_manual(values = c("grey30","salmon")) + 
  scale_color_manual(values = c("grey30","salmon")) + 
    labs(title = paste("PLAP6 (",PLAP6,")",sep=""))
  assign(paste("gcounts_",ig,sep=""),gcounts)

  gcounts <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
  geom_boxplot(fatten = 1) +
  scale_fill_manual(values = group.colors) +
  scale_y_continuous(trans = "log2")
gcounts
# ggsave(paste(outdir,"graphs/2023_11_counts_tfs_aio_log2.pdf",sep="/"), plot = gcounts,
#   width = 18,
#   height = 10)

goi_phot <- c("CHR1", "CHR2", "PCRY1", "ACRY1", "DCRY1", "DCRY2", "PHOT1", "UVR8", "HKR1") #"HKR2"

@@

rownames(anno_phot) <- anno_phot$geneSymbol
anno_phot <- anno_phot[goi_phot,]

# Get Count data
i <- 1
l <- length(goi_phot)
all_counts <- {}
for (i in 1:l){
  d <-  plotCounts(dds, gene=anno_phot[i,"gene_id"], intgroup=c("condition","genotype","treatment"), main=anno_phot[i,"id.symbol"],returnData=TRUE)
  d$Gene <- rep(anno_phot[i,"id.symbol"],length(rownames(d)))
  d$sample <- rownames(d)
  # rownames(d) <- {}
  all_counts <- bind_rows(all_counts,d)
  }

@@

goi_tf <- c("CCM1","LCR1", "HY5", "PHOT1", "QER7", "QER4", "QER6", "ROC15", "ROC40", "ROC66", "ROC75", "ROC114", "ROC55", "CON1", "CRB1")
anno[c("Cre17.g745697","Cre13.g567250","Cre01.g043550","Cre02.g079550"),"geneSymbol"] <- c("QER4","QER6","QER7","ROC110")
anno_tf <- subset(anno,geneSymbol %in% goi_tf | gene_id %in% c("Cre17.g745697","Cre13.g567250","Cre01.g043550","Cre02.g079550"))
anno_tf[,c("gene_id", "geneSymbol","id.symbol")]
summary(anno_tf$gene_id %in% rownames(dds))
anno_tf[!anno_tf$gene_id %in% rownames(dds),]
anno_tf <- anno_tf[anno_tf$gene_id %in% rownames(dds),]

@@

goi <- anno_tf
l <- nrow(goi)
all_counts <- {}
for (i in 1:l){
  d <-  plotCounts(dds, gene=goi[i,"gene_id"], intgroup=c("condition","treatment","genotype"),col=col,main=res$SYMBOL[i],returnData=TRUE)
  d$Gene <- rep(goi[i,"geneSymbol"],length(rownames(d)))
  d$sample <- rownames(d)
  rownames(d) <- {}
  all_counts <- bind_rows(all_counts,d)
  }

all_counts$Gene
levels(all_counts$condition)
all_counts$Gene <- factor(all_counts$Gene, levels = anno_tf$geneSymbol[c(9,6,8,7,15,2,1,4,3,5,13,11,14,10,12)])
levels(all_counts$Gene)

# colours
# group.colors <- rep(anno_colors$treatment, each = 2)
group.colors <- rep(anno_colors$treatment, each = 2)[c(1:6,7,9,11)]
names(group.colors) <- colData(dds)$condition %>% levels()
anno_colors$condition <- group.colors

max_val <- 1.0*max(all_counts$count)

# Plot
gcounts <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
  geom_boxplot(fatten = 1) +
  scale_fill_manual(values = group.colors) +
  scale_y_continuous(trans = "log2")
gcounts

cols = brewer.pal(n=9,name = 'Paired')

gcounts <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition, color=condition)) +
  geom_boxplot(color="black") +
  geom_point(position=position_dodge(width=0.75), alpha=1) +
  scale_fill_manual(values=cols) +
  scale_color_manual(values=cols) +
  scale_y_continuous(trans = "log2")
gcounts

gcounts2 <- ggplot(all_counts, aes(x = condition, y = count, fill=Gene)) +
  geom_boxplot(fatten = 1) +
  scale_fill_viridis_d() +
  scale_y_continuous(trans = "log2")
gcounts2

gcounts3 <- ggplot(all_counts[all_counts$Gene=="ROC40",], aes(x = condition, y = count, fill=treatment, color=treatment)) +
  geom_boxplot(color="black") +
  geom_point(shape=21,color="black",aes(fill=treatment),position=position_dodge(width=0.75), alpha=1) +
  scale_fill_manual(values=anno_colors$treatment) +
  scale_color_manual(values=anno_colors$treatment) +
  scale_y_continuous(trans = "log2") +
  labs(title = "ROC40") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
gcounts3

gcounts4 <- ggplot(all_counts[all_counts$Gene=="ROC114",], aes(x = condition, y = count, fill=treatment, color=treatment)) +
  geom_boxplot(color="black") +
  geom_point(shape=21,color="black",aes(fill=treatment),position=position_dodge(width=0.75), alpha=1) +
  scale_fill_manual(values=anno_colors$treatment) +
  scale_color_manual(values=anno_colors$treatment) +
  scale_y_continuous(trans = "log2") +
  labs(title = "ROC114") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
gcounts4

>>>>

# 3. Results

## Plot Counts
```{r}

goi_tf <- c("CCM1","LCR1", "HY5", "QER7", "QER4", "QER6", "ROC15", "ROC40", "ROC66", "ROC75", "ROC114", "ROC55", "CON1", "CRB1")
anno[c("Cre17.g745697","Cre13.g567250","Cre01.g043550","Cre02.g079550"),"geneSymbol"] <- c("QER4","QER6","QER7","ROC110")
anno_tf <- subset(anno,geneSymbol %in% goi_tf | gene_id %in% c("Cre17.g745697","Cre13.g567250","Cre01.g043550","Cre02.g079550"))
anno_tf[,c("gene_id", "geneSymbol","id.symbol")]
summary(anno_tf$gene_id %in% rownames(dds))
anno_tf[!anno_tf$gene_id %in% rownames(dds),]
anno_tf <- anno_tf[anno_tf$gene_id %in% rownames(dds),]

goi <- anno_tf
l <- nrow(goi)
all_counts <- {}
for (i in 1:l){
  d <-  plotCounts(dds, gene=goi[i,"gene_id"], intgroup="condition", col=col,main=res$SYMBOL[i],returnData=TRUE)
  d$Gene <- rep(goi[i,"geneSymbol"],length(rownames(d)))
  d$sample <- rownames(d)
  rownames(d) <- {}
  all_counts <- bind_rows(all_counts,d)
  }

all_counts$Gene
levels(all_counts$condition)
all_counts$Gene <- factor(all_counts$Gene)
levels(all_counts$Gene)

all_counts$Gene <- factor(all_counts$Gene, levels = c("CON1", "ROC40", "CRB1" ,  "QER4" , "ROC110",  "QER7" ,"ROC114" , "ROC75", "CCM1" , "ROC55" , "LCR1",  "QER6" , "ROC66",    "ROC15")  )
levels(all_counts$Gene)

# colours
Greys <- brewer.pal(9, name="Greys")[c(7,5)]
Paired <- brewer.pal(12, name="Paired")[c(2,1,6,5)]
colours <- c(Greys,Paired)
group.colors <- colours
group.colors <- c("#525252", "#969696", "#1F78B4", "#A6CEE3", "#E31A1C", "#FB9A99")
max_val <- 1.0*max(all_counts$count)

# Plot
gcounts <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
  geom_boxplot(fatten = 1) +
  scale_fill_manual(values = group.colors) +
  scale_y_continuous(trans = "log2")
gcounts
ggsave(paste(figures,"2023_12_counts_tfs_aio_log2.pdf",sep="/"), plot = gcounts,
  width = 12,
  height = 8)

## Phot genes all-in-one
goi_phot <- c("CHR1", "CHR2", "PCRY1", "ACRY1", "DCRY1", "PHOT1", "UVR8", "HKR1") #"HKR2" "DCRY2",

@@

goi <- anno_phot
l <- nrow(goi)
all_counts <- {}
for (i in 1:l){
  d <-  plotCounts(dds, gene=goi[i,"gene_id"], intgroup="condition", col=col,main=res$SYMBOL[i],returnData=TRUE)
  d$Gene <- rep(goi[i,"geneSymbol"],length(rownames(d)))
  d$sample <- rownames(d)
  rownames(d) <- {}
  all_counts <- bind_rows(all_counts,d)
  }

all_counts$Gene
levels(all_counts$condition)
levels(all_counts$Gene)

all_counts$Gene <- factor(all_counts$Gene)
levels(all_counts$Gene)

all_counts$Gene <- factor(all_counts$Gene, levels = c("PHOT1","CHR1","CHR2","HKR1","UVR8","DCRY1","PCRY1","ACRY1")) # "DCRY2", 
levels(all_counts$Gene)

# colours
Greys <- brewer.pal(9, name="Greys")[c(7,5)]
Paired <- brewer.pal(12, name="Paired")[c(2,1,6,5)]
colours <- c(Greys,Paired)
group.colors <- colours
group.colors <- c("#525252", "#969696", "#1F78B4", "#A6CEE3", "#E31A1C", "#FB9A99")
max_val <- 1.0*max(all_counts$count)

# Plot
gcounts <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
  geom_boxplot(fatten = 1) +
  scale_fill_manual(values = group.colors) +
  scale_y_continuous(trans = "log2")
gcounts
ggsave(paste(figures,"2023_12_counts_phots_aio_log2.pdf",sep="/"), plot = gcounts,
  width = 12,
  height = 8)

# ---------------------------
# COPY & PASTE from pCRY
## Make results
```{r, eval=FALSE, include=FALSE, echo=FALSE}
outdir <- "C:/Users/kelterbs/OneDrive - Charité - Universitätsmedizin Berlin/NGS (RNA-Seq & CRISPR)/Chlamy RNA-Seq P3044"
setwd(outdir)

dds <- readRDS(paste(outdir,"dds_23design.RDS",sep="/"))
anno <- readRDS(file=paste(outdir,"anno.RDS", sep="/"))

design(dds)
resultsNames(dds)
cbind(resultsNames(dds),c(rep(0,6)))

# res_WT_D_vs.pcry_D <- results(dds, contrast = c("condition","pcry_D","WT_D"))

# Unterschiede zwischen WT und pCRY unter versch. Belichtungen
res1 <- res_WT_D_vs.pcry_D <- results(dds, contrast=c(0,0,0,1,0,0))
res1 <- res_WT_D_vs.pcry_D <- results(dds,name = "genotype_pcry_vs_WT")

res2 <- res_WT_BL_vs.pcry_BL <- results(dds, contrast=c(0,1,0,0,1,0))

res3 <- res_WT_R_vs.pcry_R <- results(dds, contrast=c(0,0,1,0,0,1))

# Unterschiede zwischen versch. Belichtungen über beide Genotypen
# D vs. (BL+R)
res4 <- res_D_vs._BL_R <-results(dds, contrast=c(0,0,2/6,2/6,1/3,1/3))

# D vs. BL
res5 <- res_BL_vs._D <- results(dds, contrast=c(0,0,1,0,0.5,0))

# D vs. R
res6 <- res_D_vs._R <- results(dds, contrast=c(0,0,0,1,0,0.5))

# WT vs. pCRY
res7 <- res_WT_vs._pcry <- results(dds, contrast=c(0,3,0,0,1,1))

## Make top lists:

topgenes <- function(res){
a <- subset(res, padj < 0.05 & baseMean > 20 & (log2FoldChange < -0.5 | log2FoldChange > 0.5))
a$id.symbol <- anno[rownames(a),"id.symbol"]
a$symbol <- anno[rownames(a),"geneSymbol"]
a <- a[order(a$baseMean, decreasing = T),]
  a$rank.bm <- seq(1:length(rownames(a)))
a <- a[order(a$padj, decreasing = F),]
  a$rank.padj <- seq(1:length(rownames(a)))
a <- a[order(abs(a$log2FoldChange), decreasing = T),]
  a$rank.l2FC <- seq(1:length(rownames(a)))
a$rank.sum <- a$rank.l2FC+a$rank.bm+a$rank.padj
  a <- a[order(a$rank.sum),]
print(a)
}

top.res1 <- topgenes(res=res1)
summary(res1)
top.res2 <- topgenes(res=res2)
summary(res2)
top.res3 <- topgenes(res=res3)
summary(res3)
top.res4 <- topgenes(res=res4)
summary(res4)
top.res5 <- topgenes(res=res5)
summary(res5)
top.res6 <- topgenes(res=res6)
summary(res6)
top.res7 <- topgenes(res=res7)
summary(res7)

top.all <- rbind(top.res1,
         top.res2,
         top.res3,
         top.res4,
         top.res5,
         top.res6,
         top.res7)
top.all <- top.all[order(top.all$rank.sum),]
head(top.all, n=185) %>% rownames() %>% unique() %>% length()
top.all <- top.all[unique(rownames(top.all)),]

# top.all <- head(top.all, n=100)

## Shrinkage
res1_shrink <- lfcShrink(dds, res=res1, type="ashr")
res2_shrink <- lfcShrink(dds, res=res2, type="ashr")
res3_shrink <- lfcShrink(dds, res=res3, type="ashr")
res4_shrink <- lfcShrink(dds, res=res4, type="ashr")
res5_shrink <- lfcShrink(dds, res=res5, type="ashr")
res6_shrink <- lfcShrink(dds, res=res6, type="ashr")
res7_shrink <- lfcShrink(dds, res=res7, type="ashr")

### combine L2FC with padj
```{r}

res3_shrink
subset(mcols(dds), id.symbol =="SELU1") %>% rownames()
res3_shrink["Cre06.g263550",]

# DataFrame with 1 row and 5 columns
#                baseMean log2FoldChange     lfcSE      pvalue        padj
#               <numeric>      <numeric> <numeric>   <numeric>   <numeric>
# Cre06.g263550   3566.14        2.90397  0.167321 1.72033e-71 1.97322e-68

plot(res3_shrink$log2FoldChange,-log(res3_shrink$padj))
res3_shrink$logpadj <- -log(res3_shrink$padj, base=10)

-log(1.38379e-06,base = 10)

x=-2.5
60/(3^2)

y=-8*x^2 + 50
y

res3_shrink$ishs <- ifelse((-20/3*(res3_shrink$log2FoldChange^2) + 60) < res3_shrink$logpadj,TRUE,FALSE)
summary(res3_shrink$ishs)
subset(res3_shrink,ishs==TRUE & logpadj< 50 & abs(log2FoldChange)<2.5)
subset(res3_shrink,ishs==TRUE & logpadj< 50)

# for log2 -/+ 5 and padj 100 
res1_shrink$ishs <- ifelse((-20/3*(res1_shrink$log2FoldChange^2) + 60) < -log(res1_shrink$padj, base=10),TRUE,FALSE)
res2_shrink$ishs <- ifelse((-20/3*(res2_shrink$log2FoldChange^2) + 60) < -log(res2_shrink$padj, base=10),TRUE,FALSE)
res3_shrink$ishs <- ifelse((-20/3*(res3_shrink$log2FoldChange^2) + 60) < -log(res3_shrink$padj, base=10),TRUE,FALSE)
res4_shrink$ishs <- ifelse((-20/3*(res4_shrink$log2FoldChange^2) + 60) < -log(res4_shrink$padj, base=10),TRUE,FALSE)
res5_shrink$ishs <- ifelse((-20/3*(res5_shrink$log2FoldChange^2) + 60) < -log(res5_shrink$padj, base=10),TRUE,FALSE)
res6_shrink$ishs <- ifelse((-20/3*(res6_shrink$log2FoldChange^2) + 60) < -log(res6_shrink$padj, base=10),TRUE,FALSE)
res7_shrink$ishs <- ifelse((-20/3*(res7_shrink$log2FoldChange^2) + 60) < -log(res7_shrink$padj, base=10),TRUE,FALSE)

```{r}
outdir
figures <- paste(outdir,"figures",sep="/")
dir.create(figures)

## Colours
```{r}
Greys <- brewer.pal(9, name="Greys")[c(7,5)]
Paired <- brewer.pal(12, name="Paired")[c(2,1,6,5)]
colours <- c(Greys,Paired)
```

## PCA (Panel B)
```{r}
library("PCAtools")
library(ggalt)

vsd <- vst(dds, blind=FALSE)
length(vsd)

dubs <- duplicated(mcols(vsd)$id.symbol)
length(dubs)

unique <- unique(mcols(vsd)$id.symbol)
length(unique)

vsd2 <- vsd[!duplicated(mcols(vsd)$id.symbol)]
rownames(vsd2) <- mcols(vsd2)$id.symbol

# colours
Greys <- brewer.pal(9, name="Greys")[c(7,5)]
Paired <- brewer.pal(12, name="Paired")[c(2,1,6,5)]
colours <- c(Greys,Paired)
group.colors <- colours
group.colors <- c("#525252", "#969696", "#1F78B4", "#A6CEE3", "#E31A1C", "#FB9A99")
names(group.colors) <- levels(colData(dds)$condition)
group.colors

plotPCA(vsd2, intgroup="condition")
vst <- assay(vsd2)
p <- pca(vst, metadata = colData(dds), removeVar = 0.1)
screeplot(p, axisLabSize = 18, titleLabSize = 22)
biplot(p, colby = "condition")
biplot(p, showLoadings = TRUE,labSize = 5, pointSize = 5, sizeLoadingsNames = 5)

biplot(p,x= "PC1", y="PC3",
       lab = p$metadata$sample,
    colby = 'condition', colkey = group.colors,
    encircle = TRUE,
    encircleFill = TRUE,
    legendPosition = 'right',
    labSize = 4)
ggsave(paste(figures,"2023_12_PCA1-3.png",sep="/"),
  width = 6,
  height = 6)

biplot(p,x= "PC3", y="PC4",
       lab = p$metadata$sample,
    colby = 'condition', colkey = group.colors,
    encircle = TRUE,
    encircleFill = TRUE,
    legendPosition = 'right')
ggsave(paste(figures,"2023_12_PCA3-4.pdf",sep="/"),
  width = 12,
  height = 12)

pairsplot(p,colby = 'condition', colkey = group.colors)
ggsave(paste(figures,"2023_12_pairsplot.pdf",sep="/"),
  width = 12,
  height = 12)

plotloadings(p,
    rangeRetain = 0.01,
    labSize = 2.0,
    title = 'Loadings plot',
    subtitle = 'PC1, PC2, PC3, PC4, PC5',
    caption = 'Top 1% variables',
    drawConnectors = TRUE)
ggsave(paste(figures,"2023_12_loadings.pdf",sep="/"),
  width = 12,
  height = 12)

anno["Cre01.g016600",]

 biplot(p,
    lab = p$metadata$condition,
    colby = 'condition', colkey = group.colors,
    hline = 0, vline = 0,
    encircle = TRUE, encircleFill = TRUE,
    legendLabSize = 16, legendIconSize = 8.0,
    legendPosition = 'right')

pairsplot(p,
    components = getComponents(p, c(1:6)),
    triangle = TRUE, trianglelabSize = 12,
    hline = 0, vline = 0,
    pointSize = 2,
    gridlines.major = FALSE, gridlines.minor = FALSE,
    colby = 'condition', colkey = group.colors,
    title = 'Pairs plot', plotaxes = FALSE,
    margingaps = unit(c(-0.01, -0.01, -0.01, -0.01), 'cm'))

## Counts
```{r}

@@

goi_tf <- c("CCM1","LCR1", "HY5", "QER7", "QER4", "QER6", "ROC15", "ROC40", "ROC66", "ROC75", "ROC114", "ROC55", "CON1", "CRB1")
anno[c("Cre17.g745697","Cre13.g567250","Cre01.g043550","Cre02.g079550"),"geneSymbol"] <- c("QER4","QER6","QER7","ROC110")
anno_tf <- subset(anno,geneSymbol %in% goi_tf | gene_id %in% c("Cre17.g745697","Cre13.g567250","Cre01.g043550","Cre02.g079550"))
anno_tf[,c("gene_id", "geneSymbol","id.symbol")]
summary(anno_tf$gene_id %in% rownames(dds))
anno_tf[!anno_tf$gene_id %in% rownames(dds),]
anno_tf <- anno_tf[anno_tf$gene_id %in% rownames(dds),]

@@

goi <- anno_tf
l <- nrow(goi)
all_counts <- {}
for (i in 1:l){
  d <-  plotCounts(dds, gene=goi[i,"gene_id"], intgroup="condition", col=col,main=res$SYMBOL[i],returnData=TRUE)
  d$Gene <- rep(goi[i,"geneSymbol"],length(rownames(d)))
  d$sample <- rownames(d)
  rownames(d) <- {}
  all_counts <- bind_rows(all_counts,d)
  }

all_counts$Gene
levels(all_counts$condition)
all_counts$Gene <- factor(all_counts$Gene)
levels(all_counts$Gene)

all_counts$Gene <- factor(all_counts$Gene, levels = c("CON1", "ROC40", "CRB1" ,  "QER4" , "ROC110",  "QER7" ,"ROC114" , "ROC75", "CCM1" , "ROC55" , "LCR1",  "QER6" , "ROC66",    "ROC15")  )
levels(all_counts$Gene)

# colours
Greys <- brewer.pal(9, name="Greys")[c(7,5)]
Paired <- brewer.pal(12, name="Paired")[c(2,1,6,5)]
colours <- c(Greys,Paired)
group.colors <- colours
group.colors <- c("#525252", "#969696", "#1F78B4", "#A6CEE3", "#E31A1C", "#FB9A99")
max_val <- 1.0*max(all_counts$count)

# Plot
gcounts <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
  geom_boxplot(fatten = 1) +
  scale_fill_manual(values = group.colors) +
  scale_y_continuous(trans = "log2")
gcounts
ggsave(paste(figures,"2023_12_counts_tfs_aio_log2.pdf",sep="/"), plot = gcounts,
  width = 12,
  height = 8)

## Phot genes all-in-one
goi_phot <- c("CHR1", "CHR2", "PCRY1", "ACRY1", "DCRY1", "PHOT1", "UVR8", "HKR1") #"HKR2" "DCRY2",

goi <- anno_phot
l <- nrow(goi)
all_counts <- {}
for (i in 1:l){
  d <-  plotCounts(dds, gene=goi[i,"gene_id"], intgroup="condition", col=col,main=res$SYMBOL[i],returnData=TRUE)
  d$Gene <- rep(goi[i,"geneSymbol"],length(rownames(d)))
  d$sample <- rownames(d)
  rownames(d) <- {}
  all_counts <- bind_rows(all_counts,d)
  }

all_counts$Gene
levels(all_counts$condition)
levels(all_counts$Gene)

all_counts$Gene <- factor(all_counts$Gene)
levels(all_counts$Gene)

all_counts$Gene <- factor(all_counts$Gene, levels = c("PHOT1","CHR1","CHR2","HKR1","UVR8","DCRY1","PCRY1","ACRY1")) # "DCRY2", 
levels(all_counts$Gene)

# colours
Greys <- brewer.pal(9, name="Greys")[c(7,5)]
Paired <- brewer.pal(12, name="Paired")[c(2,1,6,5)]
colours <- c(Greys,Paired)
group.colors <- colours
group.colors <- c("#525252", "#969696", "#1F78B4", "#A6CEE3", "#E31A1C", "#FB9A99")
max_val <- 1.0*max(all_counts$count)

# Plot
gcounts <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
  geom_boxplot(fatten = 1) +
  scale_fill_manual(values = group.colors) +
  scale_y_continuous(trans = "log2")
gcounts
ggsave(paste(figures,"2023_12_counts_phots_aio_log2.pdf",sep="/"), plot = gcounts,
  width = 12,
  height = 8)

## Volcanos
### dark
```{r}

# dark
res <- res1_shrink
n <- 'WT vs. pCRY in dark'
nf <- 'figures/Volcano_dark.png'
l <- nrow(top.res1)
pcol <- "black"
lcol <- colours[1+1]

rownames(res) <- mcols(dds)$id.symbol
res <- res[!is.na(res$padj),]
res <- res[!is.na(res$log2FoldChange),]

shape <- ifelse(abs(res$log2FoldChange) > 5, 6,
                ifelse(res$padj < 1e-100,6,16))
summary(is.na(shape))
# shape[is.na(shape)] <- 2
names(shape)[shape == 6] <- 'outliner'
names(shape)[shape == 16] <- 'in range'

res$log2FoldChange[res$log2FoldChange > 5] <- 5
res$log2FoldChange[res$log2FoldChange < -5] <- -5
res$padj[res$padj < 1e-100] <- 1e-100
summary(res$padj < 1e-100)

keyvals <- ifelse(
    res$ishs == TRUE, pcol,
      'grey')

keyvals[is.na(keyvals)] <- 'grey'
  names(keyvals)[keyvals == pcol] <- 'highly significant'
  names(keyvals)[keyvals == 'grey'] <- 'other'

evd <- EnhancedVolcano(res,
    x = 'log2FoldChange',
    y = 'padj',
    lab = rownames(res),
    selectLab  = rownames(res)[which(names(keyvals) == 'highly significant')],
    # selectLab = anno_tf$geneSymbol,
    # boxedLabels = TRUE,
    labSize = 3,
    drawConnectors = TRUE,
    boxedLabels = TRUE,
    widthConnectors = 0.5,
    colConnectors = lcol,
    pointSize = 4.0,
    max.overlaps = 10,
    colCustom = keyvals,
    xlim = c(-5.1, 5.1),
    ylim = c(0, 101),
    ylab = "Padj (-log10)",
    title = n,
    subtitle = paste("DE genes:",l),
    # sub = "SVF",
    pCutoff = 10^(-60),
    FCcutoff = 3,
   shapeCustom =shape,
    # pointSize = c(ifelse(rownames(res_WT_D_vs.WT_BL) %in% rownames(top_WT_BL_vs.pcry_BL), 8, 1)),
    legendLabels=c('Not sig.','|L2F| > 2.5','p-adj < 0.05',
                   'p-adj & L2F'),
    legendPosition = 'bottom',
    col=c('grey', pcol, pcol, pcol)
    )
evd

### blue
```{r}

# blue
res <- res2_shrink
n <- 'WT vs. pCRY in blue'
nf <- 'figures/Volcano_blue.png'
l <- nrow(top.res2)
pcol <- colours[3]
lcol <- colours[3+1]

rownames(res) <- mcols(dds)$id.symbol
res <- res[!is.na(res$padj),]
res <- res[!is.na(res$log2FoldChange),]

shape <- ifelse(abs(res$log2FoldChange) > 5, 17,
                ifelse(res$padj < 1e-100,17,16))
summary(is.na(shape))
# shape[is.na(shape)] <- 2
names(shape)[shape == 6] <- 'outliner'
names(shape)[shape == 16] <- 'in range'

res$log2FoldChange[res$log2FoldChange > 5] <- 5
res$log2FoldChange[res$log2FoldChange < -5] <- -5
res$padj[res$padj < 1e-100] <- 1e-100
summary(res$padj < 1e-100)

keyvals <- ifelse(
    res$ishs == TRUE, pcol,
      'grey')

keyvals[is.na(keyvals)] <- 'grey'
  names(keyvals)[keyvals == pcol] <- 'highly significant'
  names(keyvals)[keyvals == 'grey'] <- 'other'

evb <- EnhancedVolcano(res,
    x = 'log2FoldChange',
    y = 'padj',
    lab = rownames(res),
    selectLab  = rownames(res)[which(names(keyvals) == 'highly significant')],
    # selectLab = anno_tf$geneSymbol,
    # boxedLabels = TRUE,
    labSize = 3,
    drawConnectors = TRUE,
    boxedLabels = TRUE,
    widthConnectors = 0.5,
    colConnectors = lcol,
    pointSize = 4.0,
    max.overlaps = 10,
    colCustom = keyvals,
    xlim = c(-5.1, 5.1),
    ylim = c(0, 101),
    ylab = "Padj (-log10)",
    title = n,
    subtitle = paste("DE genes:",l),
    # sub = "SVF",
    pCutoff = 10^(-50),
    FCcutoff = 2.5,
   shapeCustom =shape,
    # pointSize = c(ifelse(rownames(res_WT_D_vs.WT_BL) %in% rownames(top_WT_BL_vs.pcry_BL), 8, 1)),
    legendLabels=c('Not sig.','|L2F| > 2.5','p-adj < 0.05',
                   'p-adj & L2F'),
    legendPosition = 'bottom',
    col=c('grey', pcol, pcol, pcol)
    )
evb

### red
```{r}

# red
res <- res3_shrink
n <- 'WT vs. pCRY in red'
nf <- 'figures/Volcano_r.png'
l <- nrow(top.res3)
pcol <- colours[5]
lcol <- colours[5+1]

rownames(res) <- mcols(dds)$id.symbol
res <- res[!is.na(res$padj),]
res <- res[!is.na(res$log2FoldChange),]

shape <- ifelse(abs(res$log2FoldChange) > 5, 6,
                ifelse(res$padj < 1e-100,6,16))
summary(is.na(shape))
# shape[is.na(shape)] <- 2
names(shape)[shape == 6] <- 'outliner'
names(shape)[shape == 16] <- 'in range'

res$log2FoldChange[res$log2FoldChange > 5] <- 5
res$log2FoldChange[res$log2FoldChange < -5] <- -5
res$padj[res$padj < 1e-100] <- 1e-100
summary(res$padj < 1e-100)

keyvals <- ifelse(
    res$ishs == TRUE, pcol,
      'grey')

keyvals[is.na(keyvals)] <- 'grey'
  names(keyvals)[keyvals == pcol] <- 'highly significant'
  names(keyvals)[keyvals == 'grey'] <- 'other'

evr <- EnhancedVolcano(res,
    x = 'log2FoldChange',
    y = 'padj',
    lab = rownames(res),
    selectLab  = rownames(res)[which(names(keyvals) == 'highly significant')],
    # selectLab = anno_tf$geneSymbol,
    # boxedLabels = TRUE,
    labSize = 3,
    drawConnectors = TRUE,
    boxedLabels = TRUE,
    widthConnectors = 0.5,
    colConnectors = lcol,
    pointSize = 4.0,
    max.overlaps = 10,
    colCustom = keyvals,
    xlim = c(-5.1, 5.1),
    ylim = c(0, 101),
    ylab = "Padj (-log10)",
    title = n,
    subtitle = paste("DE genes:",l),
    # sub = "SVF",
    pCutoff = 10^(-50),
    FCcutoff = 2.5,
   shapeCustom =shape,
    # pointSize = c(ifelse(rownames(res_WT_D_vs.WT_BL) %in% rownames(top_WT_BL_vs.pcry_BL), 8, 1)),
    legendLabels=c('Not sig.','|L2F| > 2.5','p-adj < 0.05',
                   'p-adj & L2F'),
    legendPosition = 'bottom',
    col=c('grey', pcol, pcol, pcol)
    )
evr

#ggsave(file="figures/Vulcano_rtest.pdf",
#       width = 12,
#       height = 12)

### all
```{r}

ga <- ggarrange(evb,
                evd+rremove("ylab"),
                evr+rremove("ylab"),
                ncol = 3, nrow = 1,common.legend = TRUE, legend = "bottom",
                widths = c(1.05,1,1))

ggsave(ga, file="figures/Vulcano_all.pdf",
       width = 12,
       height = 8)

# + ggplot2::coord_cartesian(xlim=c(-10, 10))
# ggplot2::coord_cartesian(xlim=c(-5, 5)

# with ggplot
ggplot(as.data.frame(resLFC_WT_BL_vs.pcry_BL),aes(log2FoldChange,-log(padj))) +
  geom_point() +
  geom_point(data = as.data.frame(resLFC_WT_BL_vs.pcry_BL[rownames(top_WT_BL_vs.pcry_BL),]), colour = "blue") + 
  geom_point(data = as.data.frame(resLFC_WT_BL_vs.pcry_BL[rocs,]), colour = "red") +
  geom_text_repel(data = as.data.frame(resLFC_WT_BL_vs.pcry_BL[rocs,]), aes(label=anno[rocs,"id.symbol"]),colour = "red",hjust=-0.5, vjust=-1) +
  xlim(-5,5)
getwd()
ggsave("graphs/Vulcano ROCs-2.pdf",
       width = 10,
       height = 6)

## enhanced vulcano
library(airway)
  library(magrittr)

 library('DESeq2')

  dds <- DESeqDataSet(airway, design = ~ cell + dex)
  dds <- DESeq(dds, betaPrior=FALSE)
  res <- results(dds,
    contrast = c('dex','trt','untrt'))
  res <- lfcShrink(dds,
    contrast = c('dex','trt','untrt'), res=res, type = 'normal')
 EnhancedVolcano(res,
    lab = rownames(res),
    x = 'log2FoldChange',
    y = 'pvalue')

## Heatmap

### plot heatmap
```{r}
library("pheatmap")
ntd <- normTransform(dds)
top.all.hm <- head(top.all, n=100)
df <- assay(ntd)[rownames(top.all.hm),]
rownames(df) <- mcols(dds)[rownames(df),"id.symbol"]

# anno_col <- as.data.frame(colData(dds)[,c("condition","treatment","genotype")])
anno_col <- data.frame(condition = dds$condition, treatment = dds$treatment)
rownames(anno_col) <- rownames(colData(dds))

# column order
anno_col <- arrange(anno_col, match(anno_col$condition, levels(anno_col$condition)))
levels(anno_col$condition)
df <- df[,rownames(anno_col)]

annot_colors=list(condition=c(Cancer="#F0978D",Healthy="#63D7DE"))
Greys <- brewer.pal(9, name="Greys")[c(7,5)]
Paired <- brewer.pal(12, name="Paired")[c(2,1,6,5)]
colours <- c(Greys,Paired)

mycolors2 <- list(genotype = c("#525252","#969696"),
     treatment = c("#525252","#1F78B4","#E31A1C"),
     condition = colours)

names(mycolors2$genotype) <- levels(anno_col$genotype)
names(mycolors2$treatment) <- levels(anno_col$treatment)
names(mycolors2$condition) <- levels(anno_col$condition)
mycolors2

xx <- pheatmap(df, cluster_rows=TRUE, show_rownames=TRUE,
         cluster_cols=FALSE,
         annotation_col=anno_col,
         color=colorRampPalette(c("navy", "white", "red"))(50),
         annotation_colors=mycolors2
         )

ggsave("figures/2023_12_Heatmap_top.png",plot=xx,
       width = 5,
       height = 15)

### Complex Heatmap
```{r}

df
genes <- anno_tf
